Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
Fixed some references

{\color {Myblue}Section~\ref{sec:defs} introduces
  the syntax and  describes a small-step operational  semantics of the
  programming language  considered in  our analysis.  We  describe our
  liveness     analysis     for      this     lazy     language     in
  Section~\ref{sec:liveness}.   Section~\ref{sec:computing}  describes
  the formulation of this liveness  as a grammar and its approximation
  to DFA. It  also contains a proof showing the  undecidability of the
  precise  liveness in  our formulation.   Section~\ref{sec:GC-scheme}
  discusses different  ways of  handling closures  and their  pros and
  cons.      We     report     our     experimental     results     in
  Section~\ref{sec:experiments} along  with some of  our observations.
  Section~\ref{sec:relatedwork}  discusses  previous work  related  to
  garbage  collection  and liveness  and  Section~\ref{sec:conclusion}
  discusses possible extensions and concludes the paper.  }
\section{The target language---syntax and semantics}
\label{sec:defs}
Figure~\ref{fig:lang-syntax} describes the syntax  of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective} where  all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation of  closures through  the $\LET$  construct.  $\LET$s  in our
language  are lazy;  in the  expression $\LET\,\,  x \leftarrow  s\,\,
\IN\,\, e$, $x$ may occur in  $s$. This enables creation of graph-like
structures in  a pure language. The  restriction of \LET\ to  a single
definition  is for  ease  of  exposition---generalization to  multiple
definitions does not add conceptual difficulties.  We further restrict
each variable in a program to  be distinct, so that no scope shadowing
occurs---this  simplifies  the  reasoning   about  the  program.   The
$\SRETURN$  and   \SIF\  expressions  and  strict   operators  trigger
evaluation of closures.

@@

\begin{table*}[ht!]
\caption{Comparing LGC times for different strategies}
\label{tab:compare-live-clo-reach-clo}
\centering
\begin{tabular}
{| l | r | r | r | r |}
\hline
Prog Name & Heap size & Num GC & Time taken using reachablity for closure  & Time taken using liveness for closure \\ \hline
\verb@nqueens(8)@ & 10000 & 183 & 0.0087s & 0.2267s\\ \hline
\verb@lambda@&10000&11&0.0010s&4.9922s\\ \hline
\verb@lcss@&10000&1&0.00008s&34.6684s\\ \hline
\end{tabular}
\end{table*}
To compare the  effect of using reachability to  copy closures instead
of liveness we take a couple  of programs from our benchmark suite and
execute       them      with       both.        As      shown       in
Table~\ref{tab:compare-live-clo-reach-clo},  the   number  of  garbage
collections in both cases are same but copying closures using liveness
takes a lot more time compared to reachability based copying.

>>>>

{\color  {Myblue}Section~\ref{sec:defs}  introduces   the  syntax  and
  describes  a small-step  operational  semantics  of the  programming
  language  considered  in our  analysis.   We  describe our  liveness
  analysis  for  this  lazy  language  in  Section~\ref{sec:liveness}.
  Section~\ref{sec:computing}  describes   the  formulation   of  this
  liveness as a grammar and its approximation to DFA. It also contains
  a proof  showing the undecidability  of the precise liveness  in our
  formulation.   Section~\ref{sec:GC-scheme} discusses  different ways
  of  handling  closures and  their  pros  and  cons.  We  report  our
  experimental  results  in Section~\ref{sec:experiments}  along  with
  some observations.  Section~\ref{sec:relatedwork} discusses previous
  work   related    to   garbage    collection   and    liveness   and
  Section~\ref{sec:conclusion}   discusses  possible   extensions  and
  concludes the paper.  }
\section{The target language---syntax and semantics}
\label{sec:defs}
Figure~\ref{fig:lang-syntax} describes the syntax  of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective} where  all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation of  closures through  the $\LET$  construct.  $\LET$s  in our
language  are lazy;  in the  expression $\LET\,\,  x \leftarrow  s\,\,
\IN\,\, e$, $x$ may occur in  $s$. This enables creation of graph-like
structures in  a pure language. The  restriction of \LET\ to  a single
definition  is for  ease  of  exposition---generalization to  multiple
definitions does not add conceptual difficulties.  We further restrict
each variable in a program to  be distinct, so that no scope shadowing
occurs---this  simplifies  the  reasoning   about  the  program.   The
$\SRETURN$  and   \SIF\  expressions  and  strict   operators  trigger
evaluation of closures.

@@

\begin{table*}[ht!]
\caption{Comparing LGC times for different strategies}
\label{tab:compare-live-clo-reach-clo}
\centering
\begin{tabular}
{| l | r | r | r | r |}
\hline
Prog Name & Heap size & Num GC & Time taken using reachablity for closure  & Time taken using liveness for closure \\ \hline
\verb@nqueens(8)@ & 10000 & 183 & 0.0087s & 0.2267s\\ \hline
\verb@lambda@&10000&11&0.0010s&4.9922s\\ \hline
\verb@lcss@&10000&1&0.00008s&34.6684s\\ \hline
\end{tabular}
\end{table*}
To compare the  effect of using reachability to  copy closures instead
of liveness we take some representative programs from our benchmark suite and
execute       them      with       both.        As      shown       in
Table~\ref{tab:compare-live-clo-reach-clo},  the   number  of  garbage
collections in both cases are same but copying closures using liveness
takes a lot more time compared to reachability based copying.
